---
title: "Gatekeeper"
author: "Miao Yu and Lauren Petrick"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import the data and transfer data format

```{r}
# data could be downloaded here: https://github.com/isglobal-brge/brgedata
load("data/exposome.RData")
load("data/metabol_serum.Rdata")

library(Biobase)
pds <- pData(metabol_serum)
fds <- fData(metabol_serum)
eds <- exprs(metabol_serum)
exposome <- exposome[match(pds$ID,exposome$ID),]
# only use numeric vairable
row <- sapply(exposome, is.numeric)
exp <- t(exposome[,row])
phenotype <- phenotype[match(pds$ID,phenotype$ID),]
```

# Gatekeeper Discovery

```{r}
# remotes::install_github('yufree/enet')
library(enet)
library(ggnetwork)
# check correlation among exposure
enet <- geteenet(exp[-1,])

# check correlation among metabolites
mnets <- getmmnet(eds)
net <- igraph::graph_from_data_frame(mnets$net,directed = F)
igraph::V(net)$membership <- igraph::components(net)$membership
n <- ggnetwork(net)
pm1 <- ggplot(n, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey50",size=0.7) +
  geom_nodes(aes(color = as.factor(membership))) +
  guides(color=F,size=F) +
  ggtitle('Serum Metabolites network') +
  theme_blank()
# +ggsave('mnec.png')
pm1

# check correlation between metabolites and exposures
re <- getemnet(exp(eds),exp[-1,])
# screen gatekeeper
gkidx <- sapply(re$data, function(x) dim(x)[2]==4)
edsgk <- eds[gkidx,]

tb <- as.data.frame.matrix(re$me) 
na <- rownames(tb)

library(network)
library(ggnetwork)
library(igraph)
allme <- re$data[sapply(re$data, dim)[2,]==4]
allmedf <- do.call(rbind,allme)
allmedfm <- allmedf[allmedf$to%in%na,]
sum(apply(table(allmedfm$from,allmedfm$to),1,sum)>1)
# 12
mname <- names(apply(table(allmedfm$from,allmedfm$to),1,sum)[apply(table(allmedfm$from,allmedfm$to),1,sum)>1])
allmedf2 <- allmedf[allmedf$from %in% mname|allmedf$to %in% mname,]

idxx <- paste(allmedf$from,allmedf$to,allmedf$direction)
df <- allmedf[!duplicated(idxx),]
df2 <- df[!df$from%in%df$from[df$from==df$to],]
df3 <- df2[df2$to%in%na,]
df4 <- df2[!df2$to%in%na,]
df5 <- df4[df4$to %in% unique(df3$from)|df4$from %in% unique(df3$from),]
df6 <- rbind.data.frame(df3,df5)
gknamez <- df3$from

net <- as.network(df6,loops = F,directed = F,matrix.type = 'edgelist')
net %e% 'direction' <- df6$direction
n <- ggnetwork(net)
n$direction <- factor(n$direction,levels = c('positive','negative'))
library(ggnetwork)
p1 <- ggplot(n, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(color = "grey50",size=1,aes(linetype = direction)) +
  geom_nodes(aes(size=2,shape=ifelse(vertex.names %in% gknamez,'square','circle'),color= ifelse(vertex.names %in% na,'black','blue'))) +
  geom_nodelabel_repel(aes(label = ifelse(vertex.names %in% gknamez,vertex.names,NA)),
                       fontface = "bold", box.padding = unit(1, "lines"))+
  guides(size = FALSE,shape=F,color=F) +
  theme_blank() 
# +ggsave('menec.png')
p1

# png('upset.png',width = 2000,height = 1200,res = 300)
tb <- as.data.frame.matrix(re$me) 
na <- rownames(tb)
tb <- tb[na%in%gknamez,]
tb <- as.data.frame(t(tb))
na <- rownames(tb)
library(UpSetR)
listinput <- list(`C16.1`=na[tb$log.C16.1==1],`C8`=na[tb$log.C8==1],`leu`=na[tb$log.Leu==1],`lysoPC.a.C26.1`=na[tb$log.lysoPC.a.C26.1==1],`lysoPC.a.C28.0`=na[tb$log.lysoPC.a.C28.0==1],`lysoPC.a.C28.1`=na[tb$log.lysoPC.a.C28.1==1],`PC.aa.C40.6`=na[tb$log.PC.aa.C40.6==1],`PC.ae.C36.3`=na[tb$log.PC.ae.C36.3==1],`PC.ae.C38.5`=na[tb$log.PC.ae.C38.5==1],`SM.C18.1`=na[tb$log.SM.C18.1], `Spermine`=na[tb$log.Spermine],`Val`=na[tb$log.Val])
upset(fromList(listinput), order.by = "freq", nsets=10,mainbar.y.label = "Associated exposures(s)")
# dev.off()
```

# Gatekeeper association with asthma

```{r}
x <- genefilter::rowttests(edsgk,factor(phenotype$hs_asthma))
xx <- genefilter::rowttests(eds,factor(phenotype$hs_asthma))
xxx <- genefilter::rowttests(exp[-1,],factor(phenotype$hs_asthma))
sum(x$p.value<0.05)
sum(xx$p.value<0.05)
sum(xxx$p.value<0.05)

df <- edsgk[rownames(edsgk) %in% rownames(x)[x$p.value<0.05],]

# png('gka.png',width = 2000,height = 1000,res = 300)
par(mar=c(4,2,4,0)+0.1,mfrow=c(1,3))
boxplot(df[1,]~phenotype$hs_asthma,xlab='',ylab='',main='C14.1',xaxt='n')
axis(1,at=c(1,2),labels = c('non-asthma','asthma'))
boxplot(df[2,]~phenotype$hs_asthma,xlab='',ylab='',main='PC.ae.C38.5',xaxt='n')
axis(1,at=c(1,2),labels = c('non-asthma','asthma'))
boxplot(df[3,]~phenotype$hs_asthma,xlab='',ylab='',main='Val',xaxt='n')
axis(1,at=c(1,2),labels = c('non-asthma','asthma'))
# dev.off()
```

# Gene

```{r eval=FALSE}
load("data/exposome.RData")
load("data/genexpr.Rdata")

library(Biobase)
pds <- pData(genexpr)
fds <- fData(genexpr)
eds <- exprs(genexpr)
exposome <- exposome[match(pds$ID,exposome$ID),]
# only use numeric vairable
row <- sapply(exposome, is.numeric)
exp <- t(exposome[,row])
phenotype <- phenotype[match(pds$ID,phenotype$ID),]
```

# Gatekeeper Discovery

```{r eval=FALSE}
# remotes::install_github('yufree/enet')
library(enet)
# check correlation among exposure
enet <- geteenet(exp[-1,])
n <- igraph::graph_from_edgelist(as.matrix(enet$net[,c(1,2)]),directed = FALSE)
plot(n,vertex.size=0.5,vertex.label=NA)

# check correlation among metabolites
mnets <- getmmnet(eds)
n <- igraph::graph_from_edgelist(as.matrix(mnets$net[,c(1,2)]),directed = FALSE)
plot(n,vertex.size=0.5,vertex.label=NA)

# check correlation between metabolites and exposures
re <- getemnet(exp(eds),exp[-1,])
# screen gatekeeper
gkidx <- sapply(re$data, function(x) dim(x)[2]==4)
edsgk <- eds[gkidx,]
```

```{r eval=FALSE}
exposome <- exposome[match(pdu$ID,exposome$ID),]
# only use numeric vairable
row <- sapply(exposome, is.numeric)
exp <- t(exposome[,row])
phenotype <- phenotype[match(pdu$ID,phenotype$ID),]
# check correlation among metabolites
mnets <- getmmnet(edu)
n <- igraph::graph_from_edgelist(as.matrix(mnets$net[,c(1,2)]),directed = FALSE)
plot(n,vertex.size=0.5,vertex.label=NA)

# check correlation between metabolites and exposures
re <- getemnet(exp(edu),exp[-1,])
# screen gatekeeper
gkidx <- sapply(re$data, function(x) dim(x)[2]==4)
edsgk <- eds[gkidx,]
write.csv(edsgk,file = 'genegk.csv')
```

278 metabolites correlation network clusters found
28026 out of 28738 metabolites have no correlation with other metabolites
169 exposures could be used for network analysis
15 exposures correlation network clusters found
91 out of 169 exposures have no correlation with other exposures
2680 peaks with multiple exposures associated
171 peaks with 6 exposures associated
1610 peaks with 1 exposures associated
700 peaks with 2 exposures associated
203 peaks with 5 exposures associated
43 peaks with 13 exposures associated
292 peaks with 4 exposures associated
65 peaks with 10 exposures associated
101 peaks with 8 exposures associated
30 peaks with 17 exposures associated
122 peaks with 7 exposures associated
481 peaks with 3 exposures associated
7 peaks with 24 exposures associated
84 peaks with 9 exposures associated
53 peaks with 12 exposures associated
52 peaks with 14 exposures associated
33 peaks with 16 exposures associated
59 peaks with 11 exposures associated
34 peaks with 15 exposures associated
4 peaks with 27 exposures associated
11 peaks with 21 exposures associated
12 peaks with 20 exposures associated
21 peaks with 18 exposures associated
1 peaks with 39 exposures associated
6 peaks with 26 exposures associated
9 peaks with 22 exposures associated
16 peaks with 44 exposures associated
7 peaks with 30 exposures associated
1 peaks with 35 exposures associated
17 peaks with 19 exposures associated
5 peaks with 28 exposures associated
13 peaks with 23 exposures associated
2 peaks with 41 exposures associated
1 peaks with 42 exposures associated
5 peaks with 34 exposures associated
3 peaks with 29 exposures associated
8 peaks with 31 exposures associated
1 peaks with 33 exposures associated
1 peaks with 36 exposures associated
3 peaks with 25 exposures associated
1 peaks with 37 exposures associated
1 peaks with 38 exposures associated
1 peaks with 46 exposures associated
257 peaks connected with other peaks and exposures
4033 peaks connected with exposures only

```{r eval=FALSE}
x <- genefilter::rowttests(edsgk,factor(phenotype$hs_asthma))
xx <- genefilter::rowttests(eds,factor(phenotype$hs_asthma))
xxx <- genefilter::rowttests(exp[-1,],factor(phenotype$hs_asthma))
sum(p.adjust(x$p.value<0.05,'BH'))
# 89
sum(p.adjust(xx$p.value<0.05,'BH'))
# 1481
sum(p.adjust(xxx$p.value<0.05,'BH'))
# 36
```

