---
title: "health_models.Rmd"
author: "Jaime Benavides"
date: "4/14/2021"
output: html_document
---
Objective:  build health models health models (1. all 8 factors, 2. prenatal factors, and if time permits 3. postnatal ones, 4. traffic for sensitivity analysis)

```{r include=FALSE}
# add folder locations
source(here::here("code/0_01_init_directory_structure.R"))

# load packages
# list packages needed:
list.of.packages = c('tidyverse', 'haven','ggplot2','GGally','caret','readr',
                     'fBasics','raster', 'nortest','MASS','extrafont', 'forestplot',
                     'ggsn','survival','grid','gridExtra', 'cowplot','reshape2',
                     'naniar','rpart','rpart.plot','pROC', 'RColorBrewer', 'broom',
                     'car','olsrr','EnvStats','janitor','tableone', 'reshape', 
                     'magrittr', 'tidyselect', 'dplyr', 'tidyr', 'stringr', 'purrr', 'tibble',
                     'pracma', 'Matrix', 'foreach', 'iterators', 'parallel', 'snow', 'doSNOW',
                     'GPfit', 'heatmaply', 'kableExtra', 'mgcv', 'glmnet')

# check if list of packages is installed. if not, it will install ones not yet installed:
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# import packages:
lapply(list.of.packages, require, character.only = TRUE)
```

## Load data

### a. Load datasets

```{r}
# load raw challenge data 
load(paste0(raw.data.folder, "exposome_v2.RData")) 
# load processed challenge data 
load(paste0(generated.data.folder, "exposome_pren_postnatal.RData")) 
# load pre- and postnatal factors 
load(paste0(generated.data.folder, "health_model_patterns.RData"))
# scores are NOT organized by cohort as they are in the knitted report HTML files, but rather by the order of the original pre and post natal scaled data matrices
pren_factors <- factors$pre$scores[]
post_factors <- factors$post$scores[]

colnames(pren_factors) <- paste0("pren_", colnames(pren_factors))
colnames(post_factors) <- paste0("post_", colnames(post_factors))
#check if same order
any(row.names(post_factors) !=row.names(pren_factors))

all_factors <- cbind(pren_factors, post_factors)
# scale factors
all_factors <- scale(all_factors, center = FALSE, scale = apply(all_factors, 2, sd, na.rm = TRUE))
```

## Confounders & Outcome

Potential Confounders: 

Child sex
Child age
Pre- pregnancy maternal bmi
Weigh of child (6-11yrs)
Heigh of child (6-11 yrs)
Maternal weight gain during pregnancy
Gestational age
Maternal age
Cohort
Maternal education
Native
Parity
Family affluence scale

Outcome:

Child Behavior Checklist (CBCL) --> hs_Gen_Tot


```{r}
# outcome 
outcomes <- phenotype %>%
  dplyr::select(ID, hs_Gen_Tot, hs_correct_raven)

# explore outcome
hist(outcomes$hs_Gen_Tot);abline(v=mean(outcomes$hs_Gen_Tot), col="red") #right skewed distribution
mean(outcomes$hs_Gen_Tot)
var(outcomes$hs_Gen_Tot) #variance greater than mean

# explore outcome
hist(outcomes$hs_correct_raven);abline(v=mean(outcomes$hs_Gen_Tot), col="red") #right skewed distribution
mean(outcomes$hs_correct_raven)
var(outcomes$hs_correct_raven) #variance greater than mean

# covariates
covariates <- dplyr::left_join(covariates[,
c("ID", "h_cohort", "hs_child_age_None", "e3_sex_None", "h_edumc_None", "h_parity_None", "h_native_None", "e3_gac_None", "h_age_None")], exposome[, c("ID", "FAS_cat_None",  "e3_asmokcigd_p_None")], by = "ID")
summary(covariates)

# bind data
outcome_conf <- full_join(outcomes, covariates, by = "ID")

dta <- data.frame(ID = as.numeric(row.names(all_factors))) %>%
  cbind(all_factors) %>%
  dplyr::left_join(outcome_conf, by = 'ID') 
```

### Factors correlation 

```{r}

cor_mat <- dta[,-1]  %>% GGally::ggcorr(., method = "pairwise.complete.obs", 
                                  limits = F, label = FALSE, label_size = 3, label_alpha = T, 
                                  size = 3, layout.exp = 1, hjust = 0.75, color = "grey50")
cor_mat

```



### Run a linear model for each exposure

```{r}
# pm2.5 
dta_expo <- cbind(dta, pren_exposome, post_exposome)
pm2p5_model <- gam(hs_Gen_Tot ~ s(hs_pm25_yr_hs_h) + 
                        h_cohort + 
                      hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None + e3_asmokcigd_p_None + 
                        h_edumc_None + h_parity_None + h_native_None + h_age_None,
                    family = "quasipoisson",   # distribution family
                    data = dta_expo)
summary(pm2p5_model)
plot(pm2p5_model)
# extract penalty

pm2p5_model$sp

plot(pm2p5_model)

# no2
no2_model <- gam(hs_Gen_Tot ~ s(hs_no2_yr_hs_h) + 
                        h_cohort + 
                      hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None + e3_asmokcigd_p_None + 
                        h_edumc_None + h_parity_None + h_native_None + h_age_None,
                    family = "quasipoisson",   # distribution family
                    data = dta_expo)
summary(no2_model)
plot(no2_model)

# building density
buildens_model <- gam(hs_Gen_Tot ~ s(hs_builtdens300_h) + 
                        h_cohort + 
                      hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None + e3_asmokcigd_p_None + 
                        h_edumc_None + h_parity_None + h_native_None + h_age_None,
                    family = "quasipoisson",   # distribution family
                    data = dta_expo)
summary(buildens_model)
plot(buildens_model)
```

### Building health model for all factors

```{r}

health_model_all_pen_sp <- gam(hs_Gen_Tot ~ s(pren_MR1) + s(pren_MR2) + s(pren_MR3) + s(pren_MR4) + 
                      s(post_MR1) + s(post_MR2) + s(post_MR3) +s(post_MR4) +
                        h_cohort + 
                      hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None + e3_asmokcigd_p_None + 
                        h_edumc_None + h_parity_None + h_native_None + h_age_None,
                    family = "quasipoisson",   # distribution family
                    data = dta)


summary(health_model_all_pen_sp)


plot(health_model_all_pen_sp)

```

### Building health model for prenatal factors

```{r}

health_model_pren_pen_sp <- gam(hs_Gen_Tot ~ s(pren_MR1) + s(pren_MR2) + s(pren_MR3) + s(pren_MR4) + 
                          hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None +  e3_asmokcigd_p_None + 
                                                   h_cohort + 
                          h_edumc_None + h_parity_None + h_native_None + h_age_None,
                        family = "quasipoisson",   # distribution family
                        data = dta)


summary(health_model_pren_pen_sp)
plot(health_model_pren_pen_sp)

```

### Building health model for postnatal factors

```{r}

health_model_post_pen_sp <- gam(hs_Gen_Tot ~ s(post_MR1) + s(post_MR2) + s(post_MR3) + s(post_MR4) + 
                           hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None + e3_asmokcigd_p_None +  
                                                   h_cohort + 
                           h_edumc_None + h_parity_None + h_native_None + h_age_None,
                         family = "quasipoisson",   # distribution family
                         data = dta)


summary(health_model_post_pen_sp)
plot(health_model_post_pen_sp)



```

### Building health model for all factors (correct raven)

```{r}

health_model_all_pen_sp <- gam(hs_correct_raven ~ s(pren_MR1) + s(pren_MR2) + s(pren_MR3) + s(pren_MR4) + 
                      s(post_MR1) + s(post_MR2) + s(post_MR3) +s(post_MR4) +
                        h_cohort + 
                      hs_child_age_None + e3_sex_None + e3_gac_None + FAS_cat_None + e3_asmokcigd_p_None + 
                        h_edumc_None + h_parity_None + h_native_None + h_age_None,
                    family = "quasipoisson",   # distribution family
                    data = dta)


summary(health_model_all_pen_sp)


plot(health_model_all_pen_sp)

```
