---
title: "Lasso model"
author: "Yanelli Nunez"
date: "4/11/2021"
output:
 html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Objective: 

```{r include=FALSE}
# add folder locations
source(here::here("code/0_01_init_directory_structure.R"))

# load packages
# list packages needed:
list.of.packages = c('tidyverse', 'haven','ggplot2','GGally','caret','readr',
                     'fBasics','raster', 'nortest','MASS','extrafont', 'forestplot',
                     'ggsn','survival','grid','gridExtra', 'cowplot','reshape2',
                     'naniar','rpart','rpart.plot','pROC', 'RColorBrewer', 'broom',
                     'car','olsrr','EnvStats','janitor','tableone', 'reshape', 
                     'magrittr', 'tidyselect', 'dplyr', 'tidyr', 'stringr', 'purrr', 'tibble',
                     'pracma', 'Matrix', 'foreach', 'iterators', 'parallel', 'snow', 'doSNOW',
                     'GPfit', 'heatmaply', 'kableExtra', 'mgcv', 'glmnet')

# check if list of packages is installed. if not, it will install ones not yet installed:
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# import packages:
lapply(list.of.packages, require, character.only = TRUE)
```


## 1. Load data

```{r}
load(paste0(generated.data.folder, "exposome_pren_postnatal.RData")) 
load(paste0(generated.data.folder, "health_model_pcp.RData")) 

prenatal <- as.data.frame(pcp_outs$pre$S) %>%
  clean_names() %>%
 rename_all( ~ paste0("pre_", .x))

postnatal <- as.data.frame(pcp_outs$post$S) %>%
  clean_names() %>%
 rename_all( ~ paste0("post_", .x))


load(paste0(generated.data.folder, "health_model_patterns.RData"))
# scores are NOT organized by cohort as they are in the knitted report HTML files, but rather by the order of the original pre and post natal scaled data matrices


pren_factors <- as.data.frame(factors$pre$scores[]) %>%
 rename_all( ~ paste0("f_pre_", .x))

post_factos <- as.data.frame(factors$post$scores[]) %>%
 rename_all( ~ paste0("f_post_", .x))

factors <- cbind(pren_factors, post_factos) %>%
    rownames_to_column(., "ID") %>%
  mutate(ID = as.integer(ID))


load(paste0(raw.data.folder, "exposome_v2.RData")) 

extra_cov <- exposome %>%
dplyr::select( ID, FAS_cat_None,e3_asmokcigd_p_None)

```

## 2. Bind, covariants, s matrix, and outcome

Potential Confounders: 
Child sex
Child age
Pre- pregnancy maternal bmi
Weigh of child (6-11yrs)
Heigh of child (6-11 yrs)
Maternal weight gain during pregnancy
Gestational age
Maternal age
Cohort
Maternal education
Native
Parity

Outcome:
Child Behavior Checklist (CBCL) --> hs_Gen_Tot

```{r}

# PCP Sparse matrix
s.matrix <- cbind(prenatal, postnatal) %>%
 # mutate_all( ~(scale(., center = FALSE, scale = sd(., na.rm = TRUE)))) %>%
    rownames_to_column(., "ID") %>%
  mutate(ID = as.integer(ID))

# PCP factors 


# Outcome 
cbcl <- phenotype %>%
  dplyr::select(ID, hs_Gen_Tot)


hist(cbcl$hs_Gen_Tot);abline(v=mean(cbcl$hs_Gen_Tot), col="red") #right skewed distribution
mean(cbcl$hs_Gen_Tot)
var(cbcl$hs_Gen_Tot) #variance greater than mean


# Bind S matrix with outcome
data_full <- full_join(cbcl, s.matrix, by = "ID") %>% 
  full_join(., covariates, by = "ID") %>%
  dplyr::select(-hs_c_weight_None, -hs_c_height_None, -hs_wgtgain_None, -e3_yearbir_None, -h_mbmi_None) %>% 
  full_join(., extra_cov, by = "ID") %>%
 full_join(., factors, by = "ID") %>%
  dplyr::select(-ID) %>%
  clean_names()
 


#summary(data_full)


```


## 3. Data Prep for model

```{r}
library(glmnet)
# Create a matrix of predictors as x
x = model.matrix(hs_gen_tot~ ., data_full)[,-1]

dim(x)
colnames(x)

# Extract outcome vector
y = data_full$hs_gen_tot

# Define variables to penalize and not penalize
is_penalized = c(rep(1, ncol(x[,1:52])), rep(0, ncol(x[,53:78])))
```

## 4. Glmnet poisson 

Notes:  

- glmnet() fits the model for 100 values of lambda
- The function glmnet returns a sequence of models for the users to choose from
- Thus we use cv.glmnet() to identify the best lambda value using cross validation which we then specify in the glmnet() function

Info about fitting a quasi poisson or negative binomial using glmnet:
https://glmnet.stanford.edu/articles/glmnetFamily.html


```{r}

set.seed(10)

#cross validation
cv.glmnet = cv.glmnet(x, y, 
                      alpha = 1,
                      penalty.factor = is_penalized,
                      family = "poisson",
                      standardize = TRUE,
                      #trace.it = TRUE,
                      n.lambda = 200, max.iter = 20000,
                      seed = 1987)
cv.glmnet
plot(cv.glmnet)

best_lambda = cv.glmnet$lambda.min # lambda.min is the value of ðœ† that gives minimum mean cross-validated error

best_lambda

# run lasso with best lambda
lasso.glmnet = glmnet(x, y, 
                   penalty.factor = is_penalized,
                   alpha = 1, 
                   lambda = best_lambda,
                   family = "poisson",
                   standardize = TRUE,
                   seed = 1987)

coef_lasso = coef(lasso.glmnet)
coef_lasso
```

## 5. Non zero coefficients
```{r}

# Find variables that are non-zero
dimnames(coef_lasso)[[1]][which(coef_lasso != 0)]

# create a dataframe
lasso_beta = cbind(rownames(coef_lasso), as.vector(coef_lasso)) %>% 
  as.data.frame() %>% 
  rename(variable = 1, beta = 2) %>% 
  mutate(beta = as.numeric(beta)) %>% 
  filter(variable != "(Intercept)")  

# Find the number of non-zero estimates
counts_lasso <- lasso_beta %>%
  filter(beta != 0) %>%
  #group_by(chem) %>%
  summarise(number_nonzero = n())

counts_lasso

```

## 6. Visuals

```{r}

fig_data <- lasso_beta %>%
 separate(variable, c("group", "extra"), remove = FALSE) %>%
  dplyr::select(-extra) %>%
  dplyr::mutate(group = if_else(group == 'pre', "Pre-natal",
                               if_else(group == 'post', "Post-natal", 
                                       if_else(group == 'f', "Factors", "Covariates"))))

                
fig <- fig_data %>%
  filter(beta != 0) %>%
  ggplot(aes(x = beta, y = variable, color = group)) + 
  geom_point(size = 1.5, alpha = 0.7) + 
  #facet_wrap(~chem) +
  geom_vline(xintercept = 0) +
  theme_minimal() +
  ggtitle("Coefficients from Lasso for pre-, post-natal and covariates")

ggplotly(fig, tooltip = "beta")


#fig2 <- fig_data %>%
#  filter(beta != 0,
#         group != "covariates") %>%
#  ggplot(aes(x = beta, y = variable, color = group)) + 
#  geom_point(size = 1.5, alpha = 0.7) + 
#  #facet_wrap(~chem) +
#  geom_vline(xintercept = 0) +
#  theme_minimal() +
#  ggtitle("Coefficients from Lasso for pre- and post-natal exposure variables")

#ggplotly(fig2, tooltip = "beta")
```

