---
title: "PCP applied Exposome data"
author: "Lawrence Chillrud (lgc2139@cumc.columbia.edu)"
date: "4/5/2021"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
fig_width: 12
fig_height: 10
params:
  algo: "Root PCP"
  dataset: "Exposome"
  subname: ""
  parameters: "Default"
  convergence: ""
  L: NULL
  S: NULL
  rowvar_name: "ID"
  rowvar: NULL
  colnames: NULL
  colgroupings: NULL
  rowgroupings: NULL
  rowgroupname: "cohort"
  scale_flag: FALSE
  ranktol: 1e-04
  sparsitytol: 1e-04
  pcs: NULL
  run_nmf: TRUE
  fa_method: "varimax"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	fig.height = 10,
	fig.width = 12,
	warning = FALSE
)

# for PCP
library(PCPhelpers) # this file was written with version 0.3.5
library(pcpr) # this file was written with version 1.0.0

# for PCA
library(ggfortify)

# for FA
library(psych)
library(GPArotation)
library(ggrepel)

# for NNMF
library(NMF)

# for data visualization
library(ggplot2)
library(heatmaply)
library(kableExtra)

# for data wrangling
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(purrr)
```

## `r paste(params$algo, "with", params$parameters, "parameters")`
```{r}
if (params$fa_method == "varimax") {
  fam <- "Orthogonal"
} else {
  fam <- "Oblique"
}

if (is.null(params$colnames)) {
  cn <- colnames(params$mat)
} else {
  cn <- params$colnames
}

if (!is.null(params$colgroupings)) {
  cng <- data.frame("Chemical" = cn, "Grouping" = params$colgroupings)
} else {
  cng <- data.frame("Chemical" = cn)
}

if (is.null(params$rowgroupings)) {
  rgs <- rep(1, nrow(params$mat))
} else {
  rgs <- params$rowgroupings
}

if (is.null(params$rowvar)) {
  rv <- 1:nrow(params$mat)
} else {
  rv <- params$rowvar
}

plot.title <- paste(params$dataset, "data", paste0("(", params$subname, ")"))
```

```{r, init}
n <- nrow(params$L)
m <- ncol(params$L)

L.org <- params$L %>% 
  as_tibble() %>% 
  mutate(!!params$rowgroupname := params$rowgroupings) %>% 
  mutate(!!params$rowvar_name := rv) %>%
  arrange(!!sym(params$rowgroupname)) %>% 
  as.data.frame()

L.df <- L.org %>% 
  pivot_longer(cols = cn, names_to = "Chemical") %>%
  right_join(., cng, by = "Chemical")

rowlabs <- params$rowvar
cohorts <- L.org[[params$rowgroupname]] %>% as.factor()
L.org <- L.org %>% select(-!!params$rowgroupname, -!!params$rowvar_name)

L.rank <- Matrix::rankMatrix(params$L, tol = params$ranktol)

if (is.null(params$pcs)) {
  pcs <- paste0("PC", 1:L.rank)
} else {
  pcs <- params$pcs
}
  
L.eda <- eda(L.org, pcs = pcs, cor_lbl = T, scale_flag = params$scale_flag, colgroups = cng, rowgroups = cohorts)


S.org <- params$S %>% 
  as_tibble() %>% 
  mutate(!!params$rowgroupname := params$rowgroupings) %>% 
  mutate(!!params$rowvar_name := rv) %>%
  arrange(!!sym(params$rowgroupname)) %>% 
  as.data.frame()

S.df <- S.org %>% 
  pivot_longer(cols = cn, names_to = "Chemical") %>%
  right_join(., cng, by = "Chemical")

S.org <- S.org %>% select(-!!params$rowgroupname, -!!params$rowvar_name)

S.sparsity <- sparsity(params$S, tol = params$sparsitytol)

summary <- data.frame("Dataset" = plot.title, "Algorithm" = params$algo, "Parameters" = params$parameters, "Convergence" = params$convergence, "L.rank" = L.rank, "S.sparsity" = S.sparsity)

summary %>% kbl(caption = "Summary") %>% kable_classic(html_font = "Cambria", position = "center") %>% kable_styling(full_width = F, "hover") %>% footnote(symbol = c("Scaled the data matrix by column"))

```

## L matrix

```{r, L mat}
######## HEATMAP ######## 
L.hm <- heatmaply(L.org, Colv = F, Rowv = F,
                    ylab = params$rowvar_name, labRow = as.character(rowlabs), 
                    cexRow = 100, row_side_colors = data.frame("cohort" = cohorts),
                    col_side_colors = data.frame("exposure family" = as.factor(params$colgroupings)),
                    showticklabels = c(T, F), main = "L matrix", 
                    label_names = c(params$rowvar_name, "exposure", "value")
                    )

L.hm

######## LINE GRAPHS ######## 
L.line <- L.df %>% 
  rowid_to_column() %>%
  ggplot(aes(x = rowid, y = value, color = !!sym(params$rowgroupname))) + 
  geom_line() + 
  facet_wrap(~Chemical) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) + ggtitle("L matrix by column")

L.line

######## TRENDS ########
L.trends <- L.df %>% 
  rowid_to_column() %>%
  ggplot(aes(x = rowid, y = value, color = !!sym(params$rowgroupname))) + 
  geom_smooth() + 
  facet_wrap(~Chemical) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) + ggtitle("Main trends in the L matrix")

L.trends

######## VIOLIN PLOTS ######## 
L.violins <- L.df %>% 
  ggplot(aes(x = Chemical, y = value, color = !!sym(params$rowgroupname))) + 
  geom_boxplot() + ggtitle("Distribution of data") + ylim(c(0, 10)) + 
  theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

L.violins

######## CORRELATION MATRIX ######## 
L.eda$cor + ggtitle("L matrix (Pearson Correlation Matrix)")
```

## S matrix

```{r, S mat}
######## HEATMAP ######## 
S.hm <- heatmaply(S.org, Colv = F, Rowv = F,
                    ylab = params$rowvar_name, labRow = as.character(rowlabs), 
                    cexRow = 100, row_side_colors = data.frame("cohort" = cohorts),
                    col_side_colors = data.frame("exposure family" = as.factor(params$colgroupings)),
                    showticklabels = c(T, F), main = plot.title, 
                    label_names = c(params$rowvar_name, "exposure", "value")
                    )
S.hm

######## LINE GRAPHS ######## 
S.line <- S.df %>% 
  rowid_to_column() %>%
  ggplot(aes(x = rowid, y = value, color = !!sym(params$rowgroupname))) + 
  geom_line() + 
  facet_wrap(~Chemical) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) + ggtitle("Raw data")

S.line

######## TRENDS ########
S.trends <- S.df %>% 
  rowid_to_column() %>%
  ggplot(aes(x = rowid, y = value, color = !!sym(params$rowgroupname))) + 
  geom_smooth() + 
  facet_wrap(~Chemical) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) + ggtitle("Main trends in the data")

S.trends

######## VIOLIN PLOTS ######## 
S.violins <- S.df %>% 
  ggplot(aes(x = Chemical, y = value, color = !!sym(params$rowgroupname))) + 
  geom_boxplot() + ggtitle("Distribution of data") + ylim(c(0, 10)) + 
  theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

S.violins 
```

## PCA on L

```{r, PCA}
L.eda$var %>% scroll_box(width = "400px", height = "400px")
L.eda$load
L.eda$biplot1
L.eda$biplot2
L.eda$biplot3
```

## Factor Analysis on L

### `r fam` Models:

`r fam` Model (Orthogonal = want factors as independent from one another as possible, get uncorrelated results, while Oblique allows for correlated factors):
```{r, FA Ortho}
# Orthogonal Model (want factors as independent from one another as possible, get uncorrelated results):
if (params$fa_method == "varimax") {
  factors <- 1:10
} else {
  factors <- L.rank
}
orthos <- factors %>% purrr::map(~fa(L.org, nfactors = ., n.obs = n, rotate = params$fa_method, scores = "regression"))

#orthos %>% walk(print, digits = 2, sort = T)

ortho_ebics <- orthos %>% map_dbl(~.$EBIC)

best_fit <- which.min(ortho_ebics)

#### Oblique Model:

#Oblique Model (want correlated factors):
# Oblique Model (want correlated factors):
#obliques <- 1:4 %>% purrr::map(~fa(L, nfactors = ., n.obs = n, rotate = "promax", scores = "regression"))
#obliques %>% purrr::walk(print, digits = 2, sort = T)
#oblique_ebics <- obliques %>% purrr::map_dbl(~.$EBIC)
```

Next we can look at the fit indices of our orthogonal models (lower EBIC = better fit). 
```{r, FA fit indices}
data.frame("Factors" = factors, "EBIC" = ortho_ebics) %>% kbl(caption = "Orthogonal Models: Fit Indices") %>% 
  kable_classic(full_width = F, html_font = "Cambria", position = "center") %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), fixed_thead = T) %>%
  row_spec(best_fit, bold = T, color = "white", background = "#D7261E") %>%
  footnote(general = "Model with the best fit shown in red.")
```

### Parameters

Looking at the `r best_fit`-factor the Orthogonal (`r params$fa_method`) model:

```{r, FA best fit}
fa_model <- orthos[[best_fit]]

print(fa_model, digits = 2)

loadings <- as.tibble(cbind(rownames(fa_model$loadings[]), fa_model$loadings[])) %>% 
  rename(Variable = V1) 

loadings <- loadings %>% mutate_at(colnames(loadings)[str_starts(colnames(loadings), "MR")], as.numeric)

loadings$Max <- colnames(loadings[, -1])[max.col(loadings[, -1], ties.method = "first")] # should be 2:5

loadings %>% kbl(caption = "Loadings") %>% kable_classic(full_width = F, html_font = "Cambria", position = "center") %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), fixed_thead = T) %>% scroll_box(width = "100%", height = "400px")

scores <- as.tibble(cbind(rownames(fa_model$scores[]), fa_model$scores[])) %>% mutate_all(as.numeric)

scores$Max <- colnames(scores)[max.col(scores, ties.method = "first")]

scores %>% kbl(caption = "Scores") %>% kable_classic(full_width = F, html_font = "Cambria", position = "center") %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), fixed_thead = T) %>% scroll_box(width = "100%", height = "400px")

fa_pats <- loadings %>% 
  select(-Max, -Variable) %>% 
  mutate_all(as.numeric)

fa_pats <- fa_pats %>% select(sort(colnames(.))) %>% as.matrix()
print_patterns(fa_pats, colgroups = cng, pat_type = "factor", n = L.rank, title = "FA factors")
```

### Factor Correlation
```{r, FA corr}
scores %>% dplyr::select(-Max) %>% corr.test() %>% print(short=FALSE)
```

### Visualize Data
```{r, FA Viz}
loadings %>% 
  ggplot(aes(x = MR1, y = MR2, col = cng$Grouping)) + 
  geom_point() + geom_label_repel(aes(label = Variable, col = cng$Grouping),
                                  box.padding   = 0.35,
                                  point.padding = 0.5,
                                  segment.color = 'grey50') + 
  theme(legend.position = "bottom") +
  labs(title = "Variable Loadings on First and Second Factors")

if ("MR3" %in% colnames(loadings)) {
  loadings %>% 
  ggplot(aes(x = MR1, y = MR3, col = cng$Grouping)) + 
  geom_point() + geom_label_repel(aes(label = Variable, col = cng$Grouping),
                                  box.padding   = 0.35,
                                  point.padding = 0.5,
                                  segment.color = 'grey50') + 
  theme(legend.position = "none") +
  labs(title = "Variable Loadings on First and Third Factors")
}
```

```{r, FA loadings}
plot_loadings <- loadings %>% select(-Max) %>% gather(key = "Factor", value = "Loading", -Variable) %>% mutate(Factor = str_replace(Factor, "MR", "Factor "))

plot_loadings %>% 
  ggplot(aes(x = Factor, y = Loading, fill = Factor)) + geom_col(position = "dodge") +
  facet_wrap(~ Variable) + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip() + geom_hline(yintercept = 0, size = 0.2) +
  labs(title = "Variable Loadings on All Factors")
```

```{r, FA scores}
scores %>% ggplot(aes(x = Max, fill = Max)) + geom_bar() +
  labs(x = "Factors", y = "Number of Individuals", title = "Number with Highest Scores per Factor") +
  theme(legend.position = "none")

scores %>% group_by(Max) %>% summarise(n())

scores %>% gather(key = "factor", value = "score", -Max) %>% dplyr::select(-Max) %>% 
  ggplot(aes(x = score)) + geom_density() + facet_grid(factor~.) # should be MR1:MR4
```


## NMF

There are many different algorithms the `NMF` package in R provides, we'll pick the `offset` algorithm. We could also pick many different ranks. We'll just pick the number of ranks that FA found: `r best_fit`.
```{r}
if (params$run_nmf) {
  num_cores <- ceiling(parallel::detectCores() / 2)
  if (sum(L.org < 0) > 0) { L.org[L.org < 0] <- 0 }
  res <- nmf(L.org, rank = best_fit, method = "offset", nrun = 30, seed = 123456, .opt = paste0('vp', num_cores))

  # only compute the scores
  s <- featureScore(res)
  summary(s)
  # compute the scores and characterize each metagene
  s <- extractFeatures(res) 
  str(s)

  W <- basis(res) # basis matrix / metagenes / contribution matrix
  H <- coef(res) # mixture coeffecient matrix / metagene expression profiles / profile matrix ie. loadings ?

  heatmaply(W, main = "NMF basis / contribution matrix (scores)", Rowv = F, Colv = F, 
            ylab = params$rowvar_name, labRow = as.character(rowlabs),
            cexRow = 100, row_side_colors = data.frame("cohort" = cohorts), showticklabels = c(T, F))

  heatmaply(H, main = "NMF profile matrix (loadings)", Rowv = F, Colv = F,
            col_side_colors = data.frame("exposure family" = as.factor(params$colgroupings)))

  loadings <- t(H) %>% as.tibble()

  print_patterns(loadings, colgroups = cng, title = "NMF loadings")
}
```
