---
title: "Exposome Raw (Scaled) Data"
author: "Lawrence Chillrud (lgc2139@cumc.columbia.edu)"
date: "4/5/2021"
output:
  html_document:
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
fig_width: 12
fig_height: 10
params:
  dataset: ""
  subname: ""
  mat: NULL
  rowvar_name: "ID"
  rowvar: NULL
  colnames: NULL
  colgroupings: NULL
  rowgroupings: NULL
  rowgroupname: "cohort"
  scale_flag: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	fig.height = 10,
	fig.width = 12,
	warning = FALSE
)

# for PCP
library(PCPhelpers) # this file was written with version 0.3.5
library(pcpr) # this file was written with version 1.0.0

# for PCA
library(ggfortify)

# for data visualization
library(ggplot2)
library(heatmaply)

# for data wrangling
library(tidyr)
library(tibble)

# for FA
library(psych)
library(GPArotation)
library(ggrepel)

# for NNMF
library(NMF)

# for data visualization
library(ggplot2)
library(heatmaply)
library(kableExtra)

# for data wrangling
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(purrr)
```
## Overview

This document hopes to explore the `r params$dataset` (`r params$subname`) dataset, so that it may be used as a reference of comparison when looking at what PCP does to the data. 

## `r params$dataset` (`r params$subname`) Data: EDA
```{r, read in data}
if (is.null(params$colnames)) {
  cn <- colnames(params$mat)
} else {
  cn <- params$colnames
}

if (!is.null(params$colgroupings)) {
  cng <- data.frame("Chemical" = cn, "Grouping" = params$colgroupings)
} else {
  cng <- data.frame("Chemical" = cn)
}

if (is.null(params$rowgroupings)) {
  rgs <- rep(1, nrow(params$mat))
} else {
  rgs <- params$rowgroupings
}

if (is.null(params$rowvar)) {
  rv <- 1:nrow(params$mat)
} else {
  rv <- params$rowvar
}

# format for plotting:
plot.title <- paste(params$dataset, "data", paste0("(", params$subname, ")"))

n <- nrow(params$mat)
m <- ncol(params$mat)

mat.org <- params$mat %>% 
  as_tibble() %>% 
  mutate(!!params$rowgroupname := params$rowgroupings) %>% 
  mutate(!!params$rowvar_name := rv) %>%
  arrange(!!sym(params$rowgroupname)) %>% 
  as.data.frame()

mat.df <- mat.org %>% 
  pivot_longer(cols = cn, names_to = "Chemical") %>%
  right_join(., cng, by = "Chemical")

rowlabs <- params$rowvar
cohorts <- mat.org[[params$rowgroupname]] %>% as.factor()
mat.org <- mat.org %>% select(-!!params$rowgroupname, -!!params$rowvar_name)
```

### The raw data matrix
```{r, heatmap}
######## HEATMAP ######## 
hm.raw <- heatmaply(mat.org, Colv = F, Rowv = F,
                    ylab = params$rowvar_name, labRow = as.character(rowlabs), 
                    cexRow = 100, row_side_colors = data.frame("cohort" = cohorts),
                    col_side_colors = data.frame("exposure family" = as.factor(params$colgroupings)),
                    showticklabels = c(T, F), main = plot.title, 
                    label_names = c(params$rowvar_name, "exposure", "value")
                    )

hm.raw
```

### The data across rows
```{r, data over time}
######## LINE GRAPHS ######## 
raw.data <- mat.df %>% 
  rowid_to_column() %>%
  ggplot(aes(x = rowid, y = value, color = !!sym(params$rowgroupname))) + 
  geom_line() + 
  facet_wrap(~Chemical) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) + ggtitle("Raw data")

######## TRENDS ########
raw.trends <- mat.df %>% 
  rowid_to_column() %>%
  ggplot(aes(x = rowid, y = value, color = !!sym(params$rowgroupname))) + 
  geom_smooth() + 
  facet_wrap(~Chemical) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank()) + ggtitle("Main trends in the data")

raw.data
raw.trends
```

### Data distribution
```{r, dist}
######## VIOLIN PLOTS ######## 
raw.violins <- mat.df %>% 
  ggplot(aes(x = Chemical, y = value, color = !!sym(params$rowgroupname))) + 
  geom_boxplot() + ggtitle("Distribution of data") + ylim(c(0, 10)) + 
  theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

raw.violins
```

### Correlation Matrix
```{r, cor mat}
mat.eda <- eda(mat.org, cor_lbl = T, scale_flag = params$scale_flag, colgroups = cng, rowgroups = cohorts)

mat.eda$cor + ggtitle(paste(plot.title, "(Pearson Correlation Matrix)"))
```

## `r plot.title` Data: PCA
```{r, pca}
mat.eda$var

mat.eda$load

mat.eda$biplot1
mat.eda$biplot2
mat.eda$biplot3
```

## `r plot.title` Data: FA

### Orthogonal Models:

Orthogonal Model (want factors as independent from one another as possible, get uncorrelated results):
```{r}
# Orthogonal Model (want factors as independent from one another as possible, get uncorrelated results):
factors <- 1:10
orthos <- factors %>% purrr::map(~fa(mat.org, nfactors = ., n.obs = n, rotate = "varimax", scores = "regression"))

#orthos %>% walk(print, digits = 2, sort = T)

ortho_ebics <- orthos %>% map_dbl(~.$EBIC)

best_fit <- which.min(ortho_ebics)

#### Oblique Model:

#Oblique Model (want correlated factors):
# Oblique Model (want correlated factors):
#obliques <- 1:4 %>% purrr::map(~fa(L, nfactors = ., n.obs = n, rotate = "promax", scores = "regression"))
#obliques %>% purrr::walk(print, digits = 2, sort = T)
#oblique_ebics <- obliques %>% purrr::map_dbl(~.$EBIC)
```

Next we can look at the fit indices of our orthogonal models (lower EBIC = better fit). 
```{r, fit indices}
data.frame("Factors" = factors, "EBIC" = ortho_ebics) %>% kbl(caption = "Orthogonal Models: Fit Indices") %>% 
  kable_classic(full_width = F, html_font = "Cambria", position = "center") %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), fixed_thead = T) %>%
  row_spec(best_fit, bold = T, color = "white", background = "#D7261E") %>%
  footnote(general = "Model with the best fit shown in red.")
```

### Parameters

Looking at the `best_fit`-factor the Orthogonal (Varimax) model:

```{r}
fa_model <- orthos[[best_fit]]

print(fa_model, digits = 2)

loadings <- as.tibble(cbind(rownames(fa_model$loadings[]), fa_model$loadings[])) %>% 
  rename(Variable = V1) 

loadings <- loadings %>% mutate_at(colnames(loadings)[str_starts(colnames(loadings), "MR")], as.numeric)

loadings$Max <- colnames(loadings[, -1])[max.col(loadings[, -1], ties.method = "first")] # should be 2:5

loadings %>% kbl(caption = "Loadings") %>% kable_classic(full_width = F, html_font = "Cambria", position = "center") %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), fixed_thead = T) %>% scroll_box(width = "100%", height = "400px")

scores <- as.tibble(cbind(rownames(fa_model$scores[]), fa_model$scores[])) %>% mutate_all(as.numeric)

scores$Max <- colnames(scores)[max.col(scores, ties.method = "first")]

scores %>% kbl(caption = "Scores") %>% kable_classic(full_width = F, html_font = "Cambria", position = "center") %>% 
  kable_styling(bootstrap_options = c("hover", "condensed"), fixed_thead = T) %>% scroll_box(width = "100%", height = "400px")

fa_pats <- loadings %>% 
  select(-Max, -Variable) %>% 
  mutate_all(as.numeric)

fa_pats <- fa_pats %>% select(sort(colnames(.))) %>% as.matrix()
print_patterns(fa_pats, colgroups = cng, pat_type = "factor", n = best_fit, title = "FA factors")
```

### Factor Correlation
```{r}
scores %>% dplyr::select(-Max) %>% corr.test() %>% print(short=FALSE)
```

### Visualize Data
```{r}
loadings %>% 
  ggplot(aes(x = MR1, y = MR2, col = cng$Grouping)) + 
  geom_point() + geom_label_repel(aes(label = Variable, col = cng$Grouping),
                                  box.padding   = 0.35,
                                  point.padding = 0.5,
                                  segment.color = 'grey50') + 
  theme(legend.position = "bottom") +
  labs(title = "Variable Loadings on First and Second Factors")

if ("MR3" %in% colnames(loadings)) {
  loadings %>% 
  ggplot(aes(x = MR1, y = MR3, col = cng$Grouping)) + 
  geom_point() + geom_label_repel(aes(label = Variable, col = cng$Grouping),
                                  box.padding   = 0.35,
                                  point.padding = 0.5,
                                  segment.color = 'grey50') + 
  theme(legend.position = "none") +
  labs(title = "Variable Loadings on First and Third Factors")
}
```

```{r}
plot_loadings <- loadings %>% select(-Max) %>% gather(key = "Factor", value = "Loading", -Variable) %>% mutate(Factor = str_replace(Factor, "MR", "Factor "))

plot_loadings %>% 
  ggplot(aes(x = Factor, y = Loading, fill = Factor)) + geom_col(position = "dodge") +
  facet_wrap(~ Variable) + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip() + geom_hline(yintercept = 0, size = 0.2) +
  labs(title = "Variable Loadings on All Factors")
```

```{r}
scores %>% ggplot(aes(x = Max, fill = Max)) + geom_bar() +
  labs(x = "Factors", y = "Number of Individuals", title = "Number with Highest Scores per Factor") +
  theme(legend.position = "none")

scores %>% group_by(Max) %>% summarise(n())

scores %>% gather(key = "factor", value = "score", -Max) %>% dplyr::select(-Max) %>% 
  ggplot(aes(x = score)) + geom_density() + facet_grid(factor~.) # should be MR1:MR4
```

## `r plot.title` Data: NMF

There are many different algorithms the `NMF` package in R provides, we'll pick the `offset` algorithm. We could also pick many different ranks. We'll just pick the number of ranks that FA found: `r best_fit`.
```{r}
num_cores <- ceiling(parallel::detectCores() / 2)
if (sum(mat.org < 0) > 0) { mat.org[mat.org < 0] <- 0 }
res <- nmf(mat.org, rank = best_fit, method = "offset", nrun = 30, seed = 123456, .opt = paste0('vp', num_cores))

# only compute the scores
s <- featureScore(res)
summary(s)
# compute the scores and characterize each metagene
s <- extractFeatures(res) 
str(s)

W <- basis(res) # basis matrix / metagenes / contribution matrix
H <- coef(res) # mixture coeffecient matrix / metagene expression profiles / profile matrix ie. loadings ?

heatmaply(W, main = "NMF basis / contribution matrix (scores)", Rowv = F, Colv = F, 
          ylab = params$rowvar_name, labRow = as.character(rowlabs),
          cexRow = 100, row_side_colors = data.frame("cohort" = cohorts), showticklabels = c(T, F))

heatmaply(H, main = "NMF profile matrix (loadings)", Rowv = F, Colv = F,
          col_side_colors = data.frame("exposure family" = as.factor(params$colgroupings)))

loadings <- t(H) %>% as.tibble()

print_patterns(loadings, colgroups = cng, title = "NMF loadings")
```
